From physical principles to ???

### 2016-04-25 2015CompleteDerivation/completeDerivation.tex
Reviewed rationale for classical assumption

\begin{rationale}
	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of the infinitesimal parts. And knowing how the parts evolve, we also know how to evolve the whole, that elements in $\mathcal{C}$.

	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for the infinitesimal parts, then one such deterministic and reversible evolution should exist. Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when it breaks down.

	The first problem is methodological. To be able to talk about the state of a part we need a physical process that is able to distinguish them. For a pool ball we can imagine marking one spot with a different color. This allows us to track it as the ball moves or collides with other balls. For an electron, instead, we do not have a way to mark a piece. In fact when two electron scatter we can't even tell which is which, let alone what part went where. The classical assumption does not hold because we do not have suitable physical processes at our disposal.

	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down for infinitesimal parts. One case where a system is sufficiently isolated to define a state is when it is "big enough". For example, the cannonball is big compared to the photons that scatter on its surface. Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and that is not going to be sufficiently isolated. The other case where Assumption \ref{detrevass} holds is where the interaction is not symmetrical. For example, the electric charge of the system under study are much larger than the charges in the environment. Therefore the electromagnetic field generated can be considered sufficiently isolated. This also breaks down: as the parts are homogeneous they have will have the same charge density and the interaction needs to be symmetrical.

	In other words: the definition of a state requires some kind of asymmetry between system under study and environment. If we say that the system is really just infinitesimal and homogeneous pieces, then the claim to that asymmetry is lost. Therefore our justification for holding the assumption is that the pieces are small compared to some scale but large compared to another. For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.

	Other conceptual problems stem from two competing requirements. One one side we want the system under study to be isolated so that we can define independent states. On the other side we do not want the system isolated so that we can define physical process on it. For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each part would essentially live in its own separate universe.

	If we assume the whole is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts. But this does not solve the problem. Let's ignore the question of how we were able to define the states of the parts. Let's also ignore that the type of forces one would use (i.e. the ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other. We simply need to note that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particle may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually is still possible, so one still ends up with a set of degrees of freedom that evolve independently, oblivious to each other.

	The same type of problem arises on the whole system when we consider its interaction to our probe or other parts of the environment. Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, $\hat{a}=f(a)$ because the system is deterministic, $\hat{b}=g(a,b)$ because the probe needs to be affected by the system, then $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map. In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems where the parts interact in any physically meaningful way.

	One final conceptual problem is with time itself: a deterministic and reversible system cannot tell time. If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes. The problem is that deterministic and reversible time evolution is equivalent to a state variable change. So, all quantities that are invariant under state variable changes are also invariant under time evolution.

	While ultimately flawed, the classical assumption can be considered valid for a great number of macroscopic system, and that is why is very useful. Understanding its shortcoming will help us later to see how the quantum assumption solves, at least partially, some of these issues.

	As a final note: we caution in automatically assuming that the classical assumption is valid for macroscopic systems. Nowhere in the assumption the spatial extent of the system is mentioned. It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, and the assumption would not hold.
\end{rationale}

>>>>

\begin{rationale}
	The idea is that time evolution specifies a map for the state space $\mathcal{S}$ of each infinitesimal part. Knowing how the parts evolve tells us how a composite state $\mathcal{c}$ evolves as well.

	Consistently with what we said in Assumption \ref{detrevass}, if we defined a state for each particle, then a deterministic and reversible evolution on such state must exist. Yet, the idea that we can assign states to infinitesimal parts should be considered only a \emph{simplifying} assumption, and it is instructive to understand when and how it breaks down.

	The first problem is methodological. To be able to talk about the states of a part we need a physical process that is able to distinguish between them. For a pool ball we can imagine marking one spot with a different color. This allows us to track it as the ball moves or collides with other balls. For an electron, instead, we do not have a way to mark a piece. In fact when two electrons scatter we can't even tell which is which, let alone what part went where. The classical assumption may not hold because we do not have suitable physical processes at our disposal.

	The second, more conceptual, problem is that the idealization that allow us to hold Assumption \ref{detrevass} break down as the parts get smaller. One case where a system is sufficiently isolated to define a state is when it is "big enough". Recall the cannonball whose motion is sufficiently unaffected by the photons that scatter off its surface. Clearly this breaks down for infinitesimal parts: as we keep dividing the material, we will find an amount that will be affected by the interaction with the environment, and therefore not sufficiently isolated. The other case where Assumption \ref{detrevass} holds is where the interaction is not symmetrical. For example, suppose the electric charge of the system under study is much larger than the charges in the environment. The electromagnetic field generated can be considered determined only by our system with good approximation. But this also breaks down: if part of the environment is made of the same homogeneous material the charge density will be the same and the interaction between particles will be symmetrical.

	In other words: the definition of a state requires some kind of asymmetry between system under study and environment. If we try to assume that the whole universe is really made of infinitesimal homogeneous parts, then the claim to that asymmetry is lost. Therefore our justification for holding the assumption has to be that the pieces are small compared to some scale but large compared to another. For example, water is a homogeneous material if we consider it made up of parts sufficiently small for the scale of the problem at hand but big enough to contain a large number of molecules.

	Other conceptual problems stem from the following two competing requirements: on one side we want the system under study to be isolated so that we can define independent states; on the other side we do not want the system isolated so that we can define physical process on it. For example, if we were to take the state of an infinitesimal part to really be its full description, with no unstated part, then each piece would evolve independently, oblivious to the other parts. Each particle would essentially live in its own separate universe.

	If we assume the whole system is deterministic and reversible but each part is not, then their evolution can depend on the states of the other parts. But this does not solve the problem entirely. Let's ignore the question of how we were able to define the states of the parts. Let's also ignore that the type of forces one would typically use (i.e. ones that conserve energy or ones with scalar and vector potentials) are precisely the ones we derive by assuming the pieces are not interacting with each other. The problem is that we can always locally separate the evolution into independent degrees of freedom. For example, the position and momentum of two particles may effect each other during the evolution, but the average and difference in position and momentum may evolve independently.\footnote{Also note that any Hamiltonian is locally isomorphic to a free particle (citation needed)} We can always find such local decomposition, and the easiest way too see that is in terms of the transported variables: they retain the original value, they clearly evolve independently. And since the pieces are infinitesimal, a local decomposition is all that is needed. While such description may be cumbersome to achieve in practice, conceptually is still possible. So, even if the particles are in principle interacting, the system is described by degrees of freedom that evolve independently, oblivious to each other.

	The same type of problem arises on the whole system when we consider its interaction to our probe or other parts of the environment. Say $a \in A$ is the state of our system and $b \in B$ the state of our probe, the future state $\hat{a}=f(a)$ because the system is deterministic, but $\hat{b}=g(a,b)$ because the probe needs to be affected by the system. This means $(a, b) \mapsto (\hat{a}, \hat{b})$ is not a reversible map. In other words: we can't expect to have deterministic and reversible evolution at all levels of decomposable systems where the parts interact in any physically meaningful way.

	One final conceptual problem is with time itself: a deterministic and reversible system cannot tell time. If it did, we would be able to find a quantity that changes through time but is invariant under state variable changes. The problem is that deterministic and reversible time evolution is equivalent to a state variable change. In fact, at a fixed time, we can choose to describe the system with the transported state variables of any past or future times. So, all quantities that are invariant under state variable changes are also invariant under time evolution.

	In the light of what we discussed, we cannot take the classical assumption to be fundamental, in the sense that we cannot take it to apply to all of physics. While ultimately flawed, the  assumption can be considered valid for a great number of macroscopic system, and that is why is very useful. Understanding its shortcoming will help us later to see how the quantum assumption solves, at least partially, some of these issues.

	As a final note: we caution in automatically thinking that the classical assumption is valid for all macroscopic systems. It is conceptually possible to have a macroscopic system where a clear independent state cannot be assigned to each part, in which case the assumption would not hold.
\end{rationale}

\subsection{Hamiltonian mechanics}
